\chapter*{Acknowledgements}
%\chapter*{Agradecimentos}

When I began envisioning and preparing for this journey, even before my official enrollment, I could not have imagined what it would be like. João from 2019 eagerly anticipated the future, picturing beauty and excitement in delving deep into a field I had long worked in and dearly loved. \\
 \textit{Oh, sweet Summer child…} \\

The path from undergraduate to now has been difficult, perilous, and an emotional rollercoaster. No one can progress from naive enthusiasm to a completed dissertation without significant support, both technical and emotional. I was fortunate to be supported by a group of people with diverse perspectives, tastes, and wisdom, who kept me afloat when things looked grim and the hurdles seemed too high. 
That said, I would like to acknowledge Professor Pedro and Ricardo.  Thank you for letting me do my thing, even when I was wrong. Having the liberty to explore my ideas and motivations are the best takeaway from this journey. Thank you! \\

To the Virtual Care team for always answering my questions and often going out of their way to help me. Feeling so well-supported was special and definitely allowed me to focus more on my research. My special thanks to Tiago, Eliana, and Pedro. \\

To my willing colleagues that were always happy to discuss my research with me and brainstorm. My special thank you to Paulo Costa, Daniela Ferreira, Rafael Vieira, Inês Ribeiro-Vaz, João Viana and Priscila Maranhão. \\

To the Outcomes Research Lab from IPOP, especially to Ana Sofia, Patricia Redondo and Ana Ferreira for providing support and being always available to discuss all of my ideas. Working with such a team is easier and reassuring. To Carlos Saez, for providing invaluable insights to my papers. Your availability to meet with me and answer even the simplest questions was priceless. To Alexandrina Cardoso, for the help with sometimes cryptic issues with data. Your drive and willingness to help were key lessons for me. \\

To my friends, especially Amarílis, André, Catarina, Diogo, Fiama, João Ribeiro, João Teixeira, Mafalda, Maria, Mariana, Miguel Maia, Miguel Ribeiro and Sofia. The possibility to vent and forget about the problems while we are together was vital in helping me stay grounded and continue pushing forward.\\

Last but not least, to my family, for always being there for me, even though they did not quite understand why I wanted this so much. \\

\vspace{10mm}
\flushleft{João}
