\chapter*{List of Publications}
% \chaptermark{REPRODUCIBILITY}



\textbf{Core Research Papers} \\
The 8 papers described below are the core structure of this thesis (6 were already published and 2 are under review). The manuscripts are listed by order of appearance in the thesis.
\\
\begin{itemize}
    \item Coutinho-Almeida, J., Rodrigues, P., \& Cruz-Correia, R. (2021). GANs for Tabular Healthcare Data Generation: A Review on Utility and Privacy. In Discovery Science (pp. 282-291). Springer International Publishing.

    \item Coutinho-Almeida, J., Cruz-Correia, R., \& Rodrigues, P. (2022). Dataset Comparison Tool: Utility and Privacy. Stud Health Technol Inform, 294, 23-27.
    

    \item (in review) Using Machine Learning Models' feature importance to assess dataset similarity 


    \item Coutinho-Almeida, J., Saez, C., Correia, R., \& Rodrigues, P. P. (2024). Development and initial validation of a data quality evaluation tool in obstetrics real-world data through HL7-FHIR interoperable Bayesian networks and expert rules. JAMIA Open, 7(3), ooae062. \url{https://doi.org/10.1093/jamiaopen/ooae062}

    
    \item Coutinho-Almeida, J., Cruz-Correia, R. J., \& Rodrigues, P. P. (2024). Evaluating distributed-learning on real-world obstetrics data: Comparing distributed, centralized and local models. Scientific Reports, 14(1), 11128. \url{https://doi.org/10.1038/s41598-024-61371-1}

    
    \item (in review) Benchmarking institutions' health outcomes with clustering methods 

    
    \item Coutinho-Almeida, J., Silva, A. S., Redondo, P., Rodrigues, P. P., \& Ferreira, A. (2024). CDK4/6 inhibitors and endocrine therapy in the treatment of metastatic breast cancer: A real-world and propensity score-adjusted comparison. Cancer Treatment and Research Communications, 40, 100818.\url{https://doi.org/10.1016/j.ctarc.2024.100818}
    
    \item Coutinho-Almeida, J., Cardoso, A., Cruz-Correia, R., \& Pereira-Rodrigues, P. (2024). Fast Healthcare Interoperability Resources-Based Support System for Predicting Delivery Type: Model Development and Evaluation Study. JMIR Formative Research, 8, e54109. \url{https://doi.org/10.2196/54109}

\end{itemize}



\textbf{Other Publications and activities}\\
In addition, during the duration of this thesis conduction, the candidate was also the author and co-author of other papers. Although these studies were not part of the thesis core structure, they were important to improve the researcher's knowledge of the field and/or to present the results to the community. They are listed below:\\

\begin{itemize}

\item Coutinho-Almeida, J., \& Cruz-Correia, R. (2022). Developing a Process Mining Tool Based on HL7. Procedia Computer Science, 196, 501-508.\\


\item Holmgren, A., Esdar, M., Hüsers, J., \& Coutinho-Almeida, J. (2023). Health Information Exchange: Understanding the Policy Landscape and Future of Data Interoperability. Yearbook of Medical Informatics, s-0043-1768719.\\



\item Costa, P., Almeida, J., Araujo, S., Alves, P., Cruz-Correia, R., Saranto, K., \& Mantas, J. (2023). Biomedical and Health Informatics Teaching in Portugal: Current Status. Heliyon, 9(3).\\

\item Gazzarata, R., Almeida, J., Lindsköld, L., Cangioli, G., Gaeta, E., Fico, G., \& Chronaki, C. E. (2024). HL7 Fast Healthcare Interoperability Resources (HL7 FHIR) in digital healthcare ecosystems for chronic disease management: Scoping review. International journal of medical informatics, 189, 105507. https://doi.org/10.1016/j.ijmedinf.2024.105507
\end{itemize}