\chapter*{Outline}

% \chaptermark{REPRODUCIBILITY}

%\doublespacing
I believe I first starting thinking about this topic during a class of the Phd Program, called exactly HEADS - Health Data Science. The topic was about the gap between the availability and the usage of health data for secondary usage and how this was a tricky question. At the time, Professor Pedro Rodrigues stated 3 big questions that were still open in the field. The first was the reusage of the data and if it was possible, ethical and feasible to use data that was created with another purpose for secondary usage -  research. The second was about to which extent routine data sources and innovations in analytical methods alleviate the need to conduct randomised clinical trials. And third, a topic related to questions of governance, privacy, and trust when routine health data are made available for research.
I spent some time on this subject, and I decided to tackle them. This decision was motivated by the fact that these subjects relate to core challenges for research with health data and define an essential research agenda for the health data science community. Furthermore, I also believe that these subjects are related to the slow adoption of innovative digital and informatics solutions in clinical settings, which often resulted in significant losses in the potential benefits of these technologies. Despite groundbreaking discoveries in the industry and academic settings, there seems to be a substantial gap in translating these advancements into practical industry applications. This realization sparked a question: how could we effectively bridge this gap and fully leverage these technologies in healthcare?

%This outline will follow these three questions.\\
%The genesis of this thesis was rooted in a profound curiosity about the potential for enhancing clinical knowledge through rapid, high-quality technological advancements. As I delved into the interplay between healthcare and technology, I noticed a prevalent issue: the data that 



%the slow adoption of innovative digital and informatics solutions in clinical settings, which often resulted in significant losses in the potential benefits of these technologies. Despite groundbreaking discoveries in academia, there seemed to be a substantial gap in translating these advancements into practical industry applications. This realization sparked a question: how could we effectively bridge this gap and fully leverage these technologies in healthcare?
To address this, the thesis is structured as follows:\\

%improve to better suit papers
Chapter \ref{chap:intro} synthesizes the aim and specific objectives of this thesis.
Chapter \ref{chap:sota} presents a brief introduction to core concepts for the thesis, such as \ac{kdd}, \ac{ebm} or privacy and ethical concerns.\\
%Chapters \ref{chap:usecase}, 4 and 5 present the main results of the different studies developed to achieve this thesis’s three main objectives. When a chapter has more than one article with conclusions, a summary is made.
In order to enable a better sctruture of the thesis, we splitted the different works acrros the data science methodology: \ac{kdd}. Focusing on the steps of data cleaning/generation, then data aquisition and analysis and finally on the usage and application of knowledge created from the data in order to have impact in real-world. With this, chapter \ref{chap:goal1} refers to work done on the generating  and preprocessing of data. Chapter \ref{chap:goal2} focus on data acquisition methods and chapter \ref{chap:goal3} focus on enabling decisions in healthcare practice based on data, through clinical research and/or CDSS. 

Chapter \ref{chap:goal1} delves into innovative methods for enhancing data quality, a critical component of the data preparation phase in KDD. This chapter presents findings from research focused on the development and utilization of synthetic data generation and automatic data quality assessment methods. The use of Generative Adversarial Networks (GANs) to create realistic, non-sensitive datasets exemplifies the synthesis of new data collection methodologies that expand data volume while protecting privacy. Furthermore, this chapter explores automatic tools designed to assess data quality, significantly reducing manual effort and enhancing the efficiency and accuracy of data analysis. Such advancements are crucial for ensuring the integrity and reliability of datasets, which form the foundation for effective data mining and subsequent knowledge discovery.

Chapter \ref{chap:goal2} assesses health data science methods under the constraints of limited data access, aligning with the data access and data mining phases of KDD. This chapter investigates the effectiveness of distributed data approaches and benchmarking strategies that enable data science techniques to operate efficiently without comprehensive data access. The focus on distributed data analysis across multiple locations supports privacy and security, particularly critical in sensitive health data scenarios. Additionally, benchmarking these methods provides a framework to evaluate their effectiveness, ensuring that the insights derived are both accurate and actionable despite limitations in data availability.

Chapter \ref{chap:goal3} explores the practical application of health data in real-time decision-making processes, particularly in clinical settings like drug evaluation and obstetrics. This chapter relates closely to the interpretation/evaluation phase of KDD, where data insights are translated into actionable knowledge. By applying causality principles and machine learning models, this research assesses the effectiveness of breast cancer drug treatments and develops clinical decision support systems (CDSS) for obstetrics. The innovative use of Interpretable Machine Learning (IML) and Inverse Probability of Treatment Weighting (IPTW) showcases how advanced data analysis techniques can influence policy and decision-making in healthcare, demonstrating the real-world applicability of KDD processes.


Chapter \ref{chap:disc}  presents the different published studies, arranged according to their connection with the traditional \ac{kdd} steps.\\

Chapter \ref{chap:conclusion} communicates the conclusion, limitations, and future work.\\

Attachments include ethical permissions and supplementary data to some papers.

%This chapter will comprise the work done during this PhD. The works developed and corresponding papers were a search for improving data usage in several steps of the \ac{kdd} process. We can see some work dedicated to leveraging data acquisition or alternatives to it, like the works depicted in section \ref{subsec:distributed}, \ref{subsec:benchmark}, \ref{subsec:similarity}, \ref{subsec:tabular} and \ref{subsec:gans}. Others will focus more on how to use the data in order to make a difference in clinical practice like the section \ref{subsec:ipop}, \ref{subsec:obs} and \ref{subsec:dq}.
